---
title: "Antibiotics"
author: "Allyson Hur"
date: 2020-
output:
  github_document
editor_options: 
  markdown: 
    wrap: 72
---

*Purpose*: Creating effective data visualizations is an *iterative*
process; very rarely will the first graph you make be the most
effective. The most effective thing you can do to be successful in this
iterative process is to *try multiple graphs* of the same data.

Furthermore, judging the effectiveness of a visual is completely
dependent on *the question you are trying to answer*. A visual that is
totally ineffective for one question may be perfect for answering a
different question.

In this challenge, you will practice *iterating* on data visualization,
and will anchor the *assessment* of your visuals using two different
questions.

*Note*: Please complete your initial visual design **alone**. Work on
both of your graphs alone, and save a version to your repo *before*
coming together with your team. This way you can all bring a diversity
of ideas to the table!

<!-- include-rubric -->

# Grading Rubric

<!-- -------------------------------------------------- -->

Unlike exercises, **challenges will be graded**. The following rubrics
define how you will be graded, both on an individual and team basis.

## Individual

<!-- ------------------------- -->

| Category | Needs Improvement | Satisfactory |
|------------------------|------------------------|------------------------|
| Effort | Some task **q**'s left unattempted | All task **q**'s attempted |
| Observed | Did not document observations, or observations incorrect | Documented correct observations based on analysis |
| Supported | Some observations not clearly supported by analysis | All observations clearly supported by analysis (table, graph, etc.) |
| Assessed | Observations include claims not supported by the data, or reflect a level of certainty not warranted by the data | Observations are appropriately qualified by the quality & relevance of the data and (in)conclusiveness of the support |
| Specified | Uses the phrase "more data are necessary" without clarification | Any statement that "more data are necessary" specifies which *specific* data are needed to answer what *specific* question |
| Code Styled | Violations of the [style guide](https://style.tidyverse.org/) hinder readability | Code sufficiently close to the [style guide](https://style.tidyverse.org/) |

## Submission

<!-- ------------------------- -->

Make sure to commit both the challenge report (`report.md` file) and
supporting files (`report_files/` folder) when you are done! Then submit
a link to Canvas. **Your Challenge submission is not complete without
all files uploaded to GitHub.**

```{r setup}
library(tidyverse)
library(ggrepel)
```

*Background*: The data[1] we study in this challenge report the
[*minimum inhibitory
concentration*](https://en.wikipedia.org/wiki/Minimum_inhibitory_concentration)
(MIC) of three drugs for different bacteria. The smaller the MIC for a
given drug and bacteria pair, the more practical the drug is for
treating that particular bacteria. An MIC value of *at most* 0.1 is
considered necessary for treating human patients.

These data report MIC values for three antibiotics---penicillin,
streptomycin, and neomycin---on 16 bacteria. Bacteria are categorized
into a genus based on a number of features, including their resistance
to antibiotics.

```{r load}
## NOTE: If you extracted all challenges to the same location,
## you shouldn't have to change this filename
filename <- "./data/antibiotics.csv"

## Load the data
df_antibiotics <- read_csv(filename)
df_antibiotics %>% knitr::kable()
```

# Visualization

<!-- -------------------------------------------------- -->

### **q1** Prototype 5 visuals

To start, construct **5 qualitatively different visualizations of the
data** `df_antibiotics`. These **cannot** be simple variations on the
same graph; for instance, if two of your visuals could be made identical
by calling `coord_flip()`, then these are *not* qualitatively different.

For all five of the visuals, you must show information on *all 16
bacteria*. For the first two visuals, you must *show all variables*.

*Hint 1*: Try working quickly on this part; come up with a bunch of
ideas, and don't fixate on any one idea for too long. You will have a
chance to refine later in this challenge.

*Hint 2*: The data `df_antibiotics` are in a *wide* format; it may be
helpful to `pivot_longer()` the data to make certain visuals easier to
construct.

```{r}
df_antibiotics %>% head
```

```{r}
df_pivot <- df_antibiotics %>% 
  pivot_longer(
    cols = c(penicillin, streptomycin, neomycin),
    names_to = "antibiotic"
  ) %>% 
  mutate(bacteria = fct_reorder(bacteria, antibiotic))


df_pivot 
```

#### Visual 1 (All variables)

In this visual you must show *all three* effectiveness values for *all
16 bacteria*. This means **it must be possible to identify each of the
16 bacteria by name.** You must also show whether or not each bacterium
is Gram positive or negative.

```{r q1.1}
df_pivot %>% 
  mutate(bacteria = fct_reorder(bacteria, antibiotic)) %>% 
  ggplot(aes(
    x = bacteria,
    y = value
  ) ) +
  geom_point(aes(color = antibiotic, shape = gram), size = 2) +
  scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

#### Visual 2 (All variables)

In this visual you must show *all three* effectiveness values for *all
16 bacteria*. This means **it must be possible to identify each of the
16 bacteria by name.** You must also show whether or not each bacterium
is Gram positive or negative.

Note that your visual must be *qualitatively different* from *all* of
your other visuals.

```{r q1.2}
df_pivot %>% 
  ggplot(aes(
    x = antibiotic,
    y = value,
    fill = bacteria
  )) +
  geom_col(position = "dodge") +
  facet_wrap(~gram, scales = "free_x") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_log10() +
  geom_hline(
    yintercept = 0.1,
    linetype = "dotted",
    color = "red"
  )
```

#### Visual 3 (Some variables)

In this visual you may show a *subset* of the variables (`penicillin`,
`streptomycin`, `neomycin`, `gram`), but you must still show *all 16
bacteria*.

Note that your visual must be *qualitatively different* from *all* of
your other visuals.

```{r q1.3}
df_antibiotics %>% 
  ggplot(aes(penicillin,streptomycin, shape = gram )) + 
  geom_point(mapping = aes(color = bacteria)) +
  scale_y_log10() +
  scale_x_log10()  +
  geom_hline(
    yintercept = 0.1,
    linetype = "dotted",
    color = "red"
  ) +
  geom_vline(
    xintercept = 0.1,
    linetype = "dotted",
    color = "red"
  )
```

#### Visual 4 (Some variables)

In this visual you may show a *subset* of the variables (`penicillin`,
`streptomycin`, `neomycin`, `gram`), but you must still show *all 16
bacteria*.

Note that your visual must be *qualitatively different* from *all* of
your other visuals.

```{r q1.4}
df_pivot %>% 
  ggplot(aes(
    x = value,
    y = antibiotic,
    fill = gram,
  ) ) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~bacteria, scales = "free_x") +
  scale_x_log10() +
  geom_vline(
    xintercept = 0.1,
    linetype = "dotted",
    color = "red"
  )
```

#### Visual 5 (Some variables)

In this visual you may show a *subset* of the variables (`penicillin`,
`streptomycin`, `neomycin`, `gram`), but you must still show *all 16
bacteria*.

Note that your visual must be *qualitatively different* from *all* of
your other visuals.

```{r q1.5}
df_antibiotics %>% 
  ggplot(aes(streptomycin, penicillin)) + 
  geom_point(mapping = aes(color = bacteria)) +
  scale_y_log10() +
  scale_x_log10() 
```

### **q2** Assess your visuals

There are **two questions** below; use your five visuals to help answer
both Guiding Questions. Note that you must also identify which of your
five visuals were most helpful in answering the questions.

*Hint 1*: It's possible that *none* of your visuals is effective in
answering the questions below. You may need to revise one or more of
your visuals to answer the questions below!

*Hint 2*: It's **highly unlikely** that the same visual is the most
effective at helping answer both guiding questions. **Use this as an
opportunity to think about why this is.**

#### Guiding Question 1

> How do the three antibiotics vary in their effectiveness against
> bacteria of different genera and Gram stain?

*Observations* - What is your response to the question above? -

Neomycin has the highest effectiveness for negative strains of bacteria,
including Diplococcus pneumonia, streptococcus hemolyticus, and
streptococcus vindans.

Penicillin has the highest effectiveness when it comes to positive
strains of bacteria, including Aerobecter aerogenes, Escherichia coli,
Klebsiella pneumoniae, proteus vularis, mycobacterium tuberculosis,
psuedomonas aeruginosa, and salmonella schottruelleri.

Streptomycin has the widest range for the effectiveness of both positive
and negative strains. Positive strains include brucella abortus,
mycobacterium tuberculosis, and pseudomonas aeruginosa while the
negative strains include diplococcus pneumonia, streptococcus
hemoolyticus, and streptococcus vindans

Overall, neomycin was effective for mostly negative strains while
penicillin was effective for only positive strains. Streptomycin has the
largest range of both positive and negative strains. In general,
penicillin requires higher concentrations compared to the other two.

\- Which of your visuals above (1 through 5) is **most effective** at
helping to answer this question? - My second visual (the bar chart with
all the information) was the most helpful in answering the question.

\- Why? - I was able to easily compare the gram strains to each of the
different bacteria Since I had them separated by the type of antibiotic,
I was able to see the effectiveness of each antibiotic against all the
different bacteria.

#### Guiding Question 2

In 1974 *Diplococcus pneumoniae* was renamed *Streptococcus pneumoniae*,
and in 1984 *Streptococcus fecalis* was renamed *Enterococcus fecalis*
[2].

> Why was *Diplococcus pneumoniae* was renamed *Streptococcus
> pneumoniae*?

*Observations* - What is your response to the question above? -
Diplococcus Pneumoniae is high in both streptomycin and neomycin, though
it's the highest in streptomycin. Other streptococcus bacteria also have
a high amount of a streptomycin, which closeley aligns with the
biological structure of Diplococcus pneumoniae.

Streptococcus fecalis has all three, but streptomycin and penicillin are
significantly higher than neomycin. The presence of penicillin doesn't
align with the other bacteria under the streptococcus genus, which could
explain that name change.

-   Which of your visuals above (1 through 5) is **most effective** at
    helping to answer this question? - Visuals 4 and 5 - Why? - In
    visual 4, I was able to compare the values of the different
    antibiotics quite easily, especially since it was separated by
    bacteria. That's how I was able to see that the streptococcus genus
    is typically high in streptomycin. Visual 5 was helpful in answering
    the second part of the question since I was able to see that the
    penicillin level of streptococcus fecalis was considerably higher
    compared to the other bacteria in the streptococcus genus.

# References

<!-- -------------------------------------------------- -->

[1] Neomycin in skin infections: A new topical antibiotic with wide
antibacterial range and rarely sensitizing. Scope. 1951;3(5):4-7.

[2] Wainer and Lysen, "That's Funny..." *American Scientist* (2009)
[link](https://www.americanscientist.org/article/thats-funny)
